---
title       : "Introduction to Leaflet"
author      : Jason A. Heppler
maketitle: true
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: material
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = F, warning = F, echo =F}
library(leaflet)
library(RColorBrewer)
library(rgdal)
library(raster)
library(ggplot2)
library(ggmap)
library(dplyr)
library(tidyr)
```

## Leaflet - Getting Started

```{r, eval=FALSE}
# install.packages("leaflet")

library(leaflet)
library(RColorBrewer)
library(ggplot2)
library(ggmap)
library(dplyr)
library(tidyr)
```

## Leaflet

Leaflet is a popular open-source JavaScript library for creating interactive maps, used by a range of sites including the *New York Times*, the *Washington Post*, GitHub, Flickr, OpenStreetMap, Mapbox, and CartoDB.

This package makes using Leaflet in R simple.

## Leaflet

Leaflet features include:

- Interactive panning/zooming
- Compose maps using:
    - Map tiles
    - Markers
    - Polygons
    - Lines
    - Popups
    - GeoJSON
  
## Leaflet

Leaflet features include:

- Create maps from the R console or RStudio
- Embed maps in RMarkdown documents and Shiny apps
- Easily render spatial objects from the `sp` package, or data frames with latitude/longitude columns
- Use map bounds and mouse events to drive Shiny app logic

## Data Prep

```{r, eval=F}
# Superfund sites CSV
calif_superfund_sites <- read.csv("./data/calif_superfund_sites.csv")
# We'll clean up the Superfund data a bit, selecting only
# the columns we want to keep.
calif_superfund_sites <- calif_superfund_sites %>% 
  select(NAME, CITY,STATE,ZIP,LONGITUDE,LATITUDE,STATUSDATE,HRS_SCORE)
# Let's filter out California sites.
calif_superfund_sites <- subset(calif_superfund_sites,
                                  STATE %in% "CA")
# Convert the date into an R friendly version.
calif_superfund_sites$date <- as.Date(calif_superfund_sites$STATUSDATE, "%m/%d/%Y")
# Lowercase the variable names.
calif_superfund_sites <- setNames(calif_superfund_sites, tolower(names(calif_superfund_sites)))
# Extract the year to a new column
library(lubridate)
calif_superfund_sites$year <- year(calif_superfund_sites$date)
```

## {.fullscreen}

<center><p style="font-size: 40px; font-family: monospaced;">https://goo.gl/tRM95l</p></center>

## Leaflet

Let's recall our process for building maps in `ggplot`.

```{r}
library(ggplot2)
library(ggmap)
map_basic <- ggplot() + 
           geom_point(data=calif_superfund_sites, 
                      aes(x=longitude, y=latitude))
```

## {.fullscreen}

```{r, echo=F, message=F, warning=F}
map_basic
```

## Leaflet

```{r}
basemap <- get_map(location = "California",
                      zoom=6,
                      maptype = "terrain",
                      source="google",
                      color="bw")
```

## Leaflet

```{r}
superfund_map <- ggmap(basemap) + 
      geom_point(data=calif_superfund_sites, aes(x=longitude, y=latitude), 
                 size = 2, alpha = 0.4, color = "#e34a33")
```

## {.fullscreen}

```{r, echo=F, warning=F,message=F}
superfund_map
```

## Leaflet

Leaflet maps are easily initiated by invoking `leaflet()`.

```{r, fig.height=4}
library(leaflet)
leaflet() %>% addTiles() %>% 
  addMarkers(lng = -123.3656, lat = 48.4284)
```

## Leaflet - Superfund Map

Let's try mapping the first thirty Superfund sites in the dataset. 

```{r, eval = T}
superfund_leaflet <- leaflet() %>% addTiles() %>%
      addCircleMarkers(data = calif_superfund_sites)
superfund_leaflet
```

## {.fullscreen}

```{r, echo = F}
superfund_leaflet
```

## Leaflet - Popups

```{r}
leaflet() %>% addTiles() %>%
      addCircleMarkers(data=calif_superfund_sites,
                       stroke = FALSE, fillOpacity = 0.5, radius=8,
                       popup=~paste("<strong>Site:</strong>", name,
                       "<br>",
                       "<strong>Date Added:</strong>", year))
```

## Leaflet - Size

```{r}
leaflet() %>% addTiles() %>%
      addCircleMarkers(data=calif_superfund_sites,
                       stroke = FALSE, fillOpacity = 0.5, radius=~hrs_score/4,
                       popup=~paste("<strong>Site:</strong>", name,
                       "<br>",
                       "<strong>Date Added:</strong>", year))
```

## Leaflet - Colors

```{r}
pal <- colorNumeric(
  palette = "Reds",
  domain = calif_superfund_sites$year
)

leaflet() %>% addTiles() %>%
      addCircleMarkers(data=calif_superfund_sites,
                       stroke = FALSE, 
                       fillColor = ~pal(year),
                       fillOpacity = 0.5, 
                       radius=~hrs_score/4,
                       popup=~paste("<strong>Site:</strong>", name,
                                    "<br>",
                                    "<strong>Date Added:</strong>", year))
```
