---
title: "Introduction to data manipulation"
author: Lincoln Mullen
ratio: "16x10"
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: false
    theme: material
    css: custom.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)

library(historydata)
library(dplyr)
library(ggplot2)
library(readr)
```

## Introduction to data manipulation with the grammar of data {.title-slide}

## Three kinds of data

- **Trees**: a branching, hierachical structure; leaves contain other leaves or any kind of data. *Examples:* directory structure of a computer; the text of a book.
- **Tables**: rectangular data with rows and columns; different columns may contain different kinds of data. *Examples:* census data.
      - **Matrices**: A special kind of rectangular data where all the columns contain only one kind of data; permits quick computation and compact storage. *Examples:* document-term matrix; network adjacency matrix.

## Tidy data

In tidy data, according to Hadley Wickham's definition, 

> Each variable forms a column.

> Each observation forms a row.

> Each type of observational unit forms a table.
 
Tidy data permits a grammar of data manipulation as well as a grammar of graphics.
 
## Tidy or not?

| state          | 1790 | 1800 | 1810 | 1820 | 1830 | 1840
|----------------|------|------|------|------|------|-------
| Virginia       | 1    | 2    | 3    | 4    | 5    | 6
| Massachusetts  | 7    | 8    | 9    | 10   | 11   | 12
| New York       | 13   | 14   | 15   | 16   | 17   | 18
| Connecticut    | 19   | 20   | 21   | 22   | 23   | 24

## Tidy or not?

| state         | year   | population
|---------------|--------|---------------
| Virginia      | 1790   | 1
| Massachusetts | 1790   | 7
| New York      | 1790   | 13
| Connecticut   | 1790   | 19
| Virginia      | 1800   | 2
| Massachusetts | 1800   | 8
| New York      | 1800   | 14
| Connecticut   | 1800   | 20

## Tidy or not?

```{r, echo=FALSE}
dijon_prices_wide %>% select(commodity, measure, `1568`, `1569`, `1570`, 
                             `1571`, `1572`, `1573`, `1574`, `1575`) %>% 
  head(8)
```

## Tidy or not?

```{r, echo=FALSE}
dijon_prices %>% select(-citation, -citation_date)
```

## Tidy or not?

| year | french_frigates | spanish_frigates | french_galleons | spanish_galleons
|------|-----------------|------------------|-----------------|-------------------
| 1810 |               92| 390              | 42              | 32
| 1811 |               95| 396              | 31              | 31
| 1812 |               90| 330              | 28              | 35
| 1813 |               86| 320              | 23              | 21 

## Grammar of data manipulation (dplyr + tidyr)

- `select()`: pick columns/variables
- `filter()`: pick rows/observations
- `arrange()`: sort by a column or set of columns
- `mutate()`: compute a new column from existing columns
- `left_join()`: connect two datasets with a common variable
- `group_by()` and `summarize()`: aggregate data in groups
- `gather()` and `spread()`: switch between wide and long data

## Folow along

Open the file `worksheets/intro-data-manipulation.Rmd`. Follow along to practice the grammar of data manipulation.

## Exercise: data manipulation worksheet

With a partner, ppen the file `worksheets/data-manipulation-worksheet.Rmd`. Work to complete the exercises.

## Exercise: working with data

### Option one: tidy other people's historical data

With a partner, go to `mappingoccupation.org` or `www.fjc.gov/history/home.nsf/page/judges.html`. Download the dataset and open it. What would that data look like in tidy format?

### Option two: tidy your own data

Begin applying the grammar of data manipulation to your own data.
