---
title: "htmlwidgets Worksheet"
author: "Jason Heppler"
date: "June 9, 2016"
output: html_document
---

## Aim of this worksheet

After completing the worksheet you should have a grasp of the various features of the [htmlwidgets](http://htmlwidgets.org) library for making interactive visualizations. You will find the [ggplot2](http://htmlwidgets.org) documentation helpful.

You are free to attempt working through this worksheet with your own data, but we encourage you for now to use the sample datasets contained in `historydata` until the final section of this worksheet.

We'll load a few necessary packages.

```{r}
library(htmlwidgets)
library(historydata)
library(dplyr)
library(tidyr)
```

First, read through the documentation: `?htmlwidgets`.

Let's start working with the `early_colleges` data.

```{r}
library(historydata)
data("early_colleges")
```

## Interactive Tables with `DT`

We'll build an interactive table of our college data.

```{r}
library(DT)
datatable(early_colleges)
```

(@) Change the default number of entries shown.

```{r}

```

(@) Try reordering the columns into a new order.

```{r}

```

## Interactive Plots with `Plotly`

Let's try working with Plotly to build an interactive chart of wholesale market prices in Dijon between 1568 to 1630. 

Load up the necessary data and package.

```{r}
library(plotly)
data("dijon_prices")
```

Read up on the documentation of `?dijon_prices` to get a sense of what's included in the dataset.

(@) Create a chart of wholesale food items measured by `pint`. Select which glyphs would work best for visualizing the data, making sure to use good labels and useful mouseover information. You may need to draw on your knowledge of `dyplyr`.

```{r}
ggiris <- qplot(Petal.Width, Sepal.Length, data = iris, color = Species)
ggplotly(ggiris)
```

(@) Give the chart some color: color the points by the `commodity`.

```{r}

```

(@) Good news! Plotly works with `ggplot2` as well! Read up [on the documentation](https://plot.ly/ggplot2) and try recreating the above chart using the `ggplot2` Plotly syntax.

## Interactive Networks with `networkD3` and `visNetwork`



## Exploring Timeseries with `dygraphs`

`dygraphs` lets us create interactive timeseries charts. These charts can be especially useful for zooming in on specific timeframes to more closely explore your data.

One tricky thing about using `dygraphs` is it expects your data to be a [timeseries object](TODO). If you're using your own data, you'll need to read up on using [xts](https://cran.rstudio.com/web/packages/xts/index.html).

ggplot
(@) Just for fun, try adding `coord_polar()` to the chart above.

```{r}

```

Neat, huh?

http://rstudio.github.io/dygraphs/index.html

## Create your own plots

Create three plots below, using any of the `historydata` datasets or your own. Your three plots should try to make some kind of humanistic observation. For each plot, include no more than three sentences explaining what you think the plot means. You should try to make each plot as informative as possible by using different geoms and including as many variables as is reasonable in each plot. Be sure to add good titles and labels.

(@) Plot 1

```{r}

```

> Explanation of plot 1.

(@) Plot 2

```{r}

```

> Explanation of plot 2.

(@) Plot 3

```{r}

```

> Explanation of plot 3.
